EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-08-14 draft.tex
Updated table
new table and figures

\begin{table}[h!tbp]
  \small
  \caption{Parameters used in the experiments.
  }
  \label{tab:params:10}
  \vspace{0.25cm}
  \centering
  \small
  \begin{tabular}{|c|c|c|c|}
    \hline
      Populations & Population Size & Generations & Maximum Iterations  \\ \hline
                 \multicolumn{4}{|c|}{10 Dimension } \\ \hline
         5        &  70             &        50   &   60                \\ \hline
         10       &  70             &        50   &   60                \\ \hline
         20       &  70             &        50   &   60                \\ \hline
                  \multicolumn{4}{|c|}{20 Dimension } \\ \hline
         5        &  70             &        50   &   60                \\ \hline
         10       &  70             &        50   &   60                \\ \hline
         20       &  70             &        50   &   60                \\ \hline

@@

%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_diff}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[tbp]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrr}
    \hline
              &         & 5 Populations    & 10 Populations \\
     Dimension& Workers & Median of rate & Median of rate  \\
    \hline
          2   & 1w      & 2.96   & 3.15   \\
              & 2w      & 1.84   & 1.72   \\
              & 4w      & 1.22   & \textbf{0.96}   \\
              & 8w      & 1.05   & \textbf{0.59}   \\
    \hline
          3   & 1w      & 6.53   & \textbf{7.76}   \\
              & 2w      & 3.64   & 3.83   \\
              & 4w      & 2.33   & 2.26   \\
              & 8w      & 2.17   &  \textbf{1.36}   \\
    \hline
          5   & 1w      & \textbf{12.33}  & 14.19  \\
              & 2w      & 6.63   & \textbf{7.41}   \\
              & 4w      & 4.25   & 4.34   \\
              & 8w      & 3.67   &  \textbf{2.41}   \\
    \hline
          10  & 1w      & \textbf{31.50}  & 38.30  \\
              & 2w      & \textbf{16.07}  & 19.31  \\
              & 4w      &  9.99   & 10.35  \\
              & 8w      & 8.97   &  \textbf{5.92}  \\
    \hline
          20  & 1w      &  116.35 & 124.51 \\
              & 2w      & \textbf{51.18}  & 63.68  \\
              & 4w      & 32.26  & 33.07  \\
              & 8w      & 28.64  & \textbf{18.25}  \\
    \hline
          40  & 1w      & \textbf{369.89} & 448.89 \\
              & 2w      & \textbf{198.43} & 234.50 \\
              & 4w      & 122.39 &  120.62 \\
              & 8w      & 122.31 & \textbf{63.35} \\
    \hline
  \end{tabular}
\end{table}

@@

%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalspersecond}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
  \label{fig:spworker:evalspersecond}
\end{figure}
%
\subsubsection{Results Regarding Speedup}
\label{sec:speedup}

@@

  \begin{tabular}{c|r|r|r|r|r|r|}
  \cline{2-7}
  \multicolumn{1}{l|}{}           & \multicolumn{3}{c|}{5 Populations}                                          & \multicolumn{3}{c|}{10 Populations}                                         \\ \hline
  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
  \multicolumn{1}{|c|}{2}         & \textbf{1.94}           & 3.12                    & 3.86                    & 1.84                    & \textbf{3.38}                    & \textbf{5.98}                    \\ \hline
  \multicolumn{1}{|c|}{3}         & \textbf{1.93}           & 2.77                    & 3.15                    & 1.90                    & \textbf{3.26}                   & \textbf{5.96}                    \\ \hline
  \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & \textbf{3.46}                    & \textbf{6.14}                    \\ \hline
  \multicolumn{1}{|c|}{10}        & \textbf{1.92}           & 3.03                    & 3.30                    & 1.90                    & \textbf{3.61}                   & \textbf{6.41}                    \\ \hline
  \multicolumn{1}{|c|}{20}        & \textbf{1.91}           & 3.14                    & 3.27                    & 1.89                    & \textbf{3.73}                   & \textbf{6.71}                    \\ \hline
  \multicolumn{1}{|c|}{40}        & \textbf{2.00}           & 2.92                    & 3.12                    & 1.95                    & \textbf{3.80}                     & \textbf{7.52}                    \\ \hline
  \end{tabular}
\end{table}
%
As expected, within the same number of populations, there is a linear speedup
when we add more workers. The speedup is marginal in the case of five
populations when increasing the number of workers above 4. The reason for this
is that at any given time, there are at least two populations in the message
queue, leaving four workers waiting. We can also notice that the spread in the
amount of work is less when using up to four workers.

\begin{table}[h!tbp]
  \small
  \caption{Speedup achieved with increasing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{lllllll}
            &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
  Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
  2         & 1w            & 2w            & {\ul 1.94}            & 6E-16           & 1.84             & 2E-16           \\
            & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
            & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
  3         & 1w            & 2w            & {\ul 1.93}            & 2E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
            & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
  5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
            & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
            & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
  10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
            & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.09}      & 1E-03           & 1.78             & 2E-16           \\
  20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
            & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.04 }     & 1E-02           & 1.80             & 2E-16           \\
  40        & 1w            & 2w            & \textbf{2.00}   & 2E-16           & 1.95             & 2E-16           \\
            & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
            & 4w            & 8w            & {\ul 1.07}            & 3E-02           & \textbf{1.98}    & 2E-16          
  \end{tabular}
  \end{table}
%
Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
one worker. When using eight workers, we achieved the maximum speed up for five
and ten populations, but interestingly, it was in dimension 40 and dimension 2 
(these values are shown in bold in the table).

>>>>

\begin{table}[h!tbp]
  \small
  \caption{Parameters used in the experiments.
  }
  \label{tab:params:10}
  \vspace{0.25cm}
  \centering
  \small
  \begin{tabular}{|c|c|c|c|}
    \hline
      Populations & Population Size & Generations & Maximum Iterations  \\ \hline
                 \multicolumn{4}{|c|}{10 Dimension } \\ \hline
         5        &  70             &        50   &   60                \\ \hline
         10       &  30             &        15   &   200                \\ \hline
         20       &  15             &        10   &   400                \\ \hline
                  \multicolumn{4}{|c|}{20 Dimension } \\ \hline
         5        &  100            &        66   &   60                \\ \hline
         10       &  60             &        15   &   200                \\ \hline
         20       &  30             &        8   &    400                \\ \hline

@@

%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{secs_per_instance}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes. The y-axis is shown in log scale.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[tbp]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrrr}
    \hline
              &         & 5 Populations    & 10 Populations & 20 Populations \\
     Dimension& Workers & Median of rate & Median of rate & Median of rate  \\
    \hline
          10  & 1w      & 22.24  & 20.48           & 24.55  \\
              & 2w      & 10.31  & 10.00           & 12.79 \\
              & 4w      &  6.23  & 4.75            &  5.70 \\
              & 8w      &  5.86  &  2.33           & 2.74 \\
              & 16w     &  5.96  &  2.33           & \textbf{1.59}\\

    \hline
          20  & 1w   & 18.36   & 65.18  & 80.31 \\
              & 2w   & 34.18   & 33.46  & 38.87 \\
              & 4w   & 19.55   & 16.58  & 18.69 \\
              & 8w   & 18.38   & 7.77   & 8.66 \\
              & 16w  & 18.36   & 7.27   & \textbf{4.57} \\
    \hline
  \end{tabular}
\end{table}

@@

%
%

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evals_per_instance}
  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
  \label{fig:spworker:evalspersecond}
\end{figure}

%

\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{eval_ratio}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
  \label{fig:spworker:evalspersecond}
\end{figure}

\subsubsection{Results Regarding Speedup}
\label{sec:speedup}

@@

  \begin{tabular}{c|r|r|r|r|}
  \hline
  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{16W}  \\ \hline
  \multicolumn{5}{| c|}{5 Populations} \\ \hline
  \multicolumn{1}{|c|}{10}        & 2.15           & 3.61                    & 3.75                    & 3.96     \\ \hline
  \multicolumn{1}{|c|}{20}        & 2.09           & 3.53                    & 3.83                    & 3.95     \\ \hline
  \multicolumn{5}{| c|}{10 Populations} \\ \hline
  \multicolumn{1}{|c|}{10}        & 1.91           & 3.82                    & 7.67                    & 8.25     \\ \hline
  \multicolumn{1}{|c|}{20}        & 1.97           & 3.97                    & 7.90                    & 8.90     \\ \hline
  \multicolumn{5}{| c|}{20 Populations} \\ \hline
  \multicolumn{1}{|c|}{10}        & 1.96           & 4.05                    & 8.29                    & 14.50     \\ \hline
  \multicolumn{1}{|c|}{20}        & 2.00           & 4.09                    & 8.18                    & 15.16     \\ \hline
  \end{tabular}
\end{table}
%
As expected, within the same number of populations, there is a linear speedup
when we add more workers. The speedup is marginal in the case of five
populations when increasing the number of workers above 4. The reason for this
is that at any given time, there are at least two populations in the message
queue, leaving four workers waiting. We can also notice that the spread in the
amount of work is less when using up to four workers.

\begin{table}[h!tbp]
  \small
  \caption{Speedup achieved with increasing the number of workers for
    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
  % Say what highlights mean - JJ
  \label{tab:speedup:test}
  \vspace{0.25cm}
  \centering
  \begin{tabular}{llllll}
  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
  & 10 & 8w & 16w & 1.05 & 0.072660 \\
  & 20 & 1w & 2w  & 2.09 & 1.51E-11 \\
  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
  & 20 & 8w & 16w & 1.03 & 0.214482 \\
  &    &    &     &      &          \\
10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
  &    &    &     &      &          \\
 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
 \end{tabular}
  \end{table}
%
Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
one worker. When using eight workers, we achieved the maximum speed up for five
and ten populations, but interestingly, it was in dimension 40 and dimension 2 
(these values are shown in bold in the table).
